\bf Primal-dual algorithm for computing Nash equilibria in sequential games with imcomplete information

### 2015-05-30 nips2015.tex
ENH: better figs

\documentclass[12pt]{article}
\usepackage{amssymb}
\usepackage{color}

%% \documentclass{article} % For LaTeX2e
%% \usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage[sc, osf]{mathpazo}  %% for nice looking formulae, etc.
\usepackage{qtree}
\usepackage{natbib}
\usepackage{graphicx}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

%% prevent figures, algorithms, etc., floating across subsections
\usepackage[section]{placeins}
\makeatletter
\AtBeginDocument{%
  \expandafter\renewcommand\expandafter\subsection\expandafter{%
    \expandafter\@fb@secFB\subsection
  }%
}
\makeatother

\usepackage{bm}
\def\B#1{\bm{#1}}
%\def\B#1{\mathbf{#1}}
\def\trans{\mathsf{T}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
 \DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\im}{im}

\newtheorem{remark}{Remark}

\def \lb {{\langle}} \def \rb {{\rangle}}
\newcommand{\fro}[1]{\|#1\|_2}
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\newcommand{\argmin}{\mathop{\mathrm{argmin}}}

\usepackage{hyperref}

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}

\providecommand{\keywords}[1]{\textbf{\textit{Index terms---}} #1}

@@

\begin{abstract}
We present a simple primal-dual algorithm with a low cost per iteration for computing Nash $\epsilon$-equilibria in two-person zero-sum sequential games with imcomplete information and perfect recall (like Texas Hold'em Poker), which converges after $\mathcal{O}(1/\epsilon)$ basic iterations
(i.e iterations involving no
%o only involving basic operations like matvec, clipping, etc., and no
calls to external first-order oracles, etc.). %% Our algorithm is applicable to a broad class of two-person zero-sum games including simultaneous games and sequential games with imcomplete information and perfect recall. The applicability to the latter kind of games is thanks to the sequen e-form representation von Stengel's \cite{von1996efficient}
Our algorithm derives from the general primal-dual algorithm developed
% by Chambolle and Pock
in \cite{chambolle2010, chambolle2014ergodic}. It should be noted that the latter primal-dual algorithm is not directly applicable to compute Nash equilibria for the aforementioned games considered here since the strategy profiles of players (polyhedra) are so complex that it is not possible
% (for example, simplexes for simultaneous games, and intersections of hyperplanes and nonnegative orthants, for sequential games)
 to effectively compute % (i.e using non-iterative procedures, etc.)
euclidean
% (metric)%
 projections (the proximal operators) thereupon.
The core of our technique is to dualize the linear constraints that form the player's strategy profile, thus obtainining an equivalent saddle-point problem which is amenable to the primal-dual scheme. Unlike the EGT (Excessive Gap Technique) recently used in \cite{hoda2010smoothing, gilpinfirst}, our algorithm has a cheap and constant cost per iteration. As an application, we apply our algorithm to solve Kuhn 3-card Poker.
%% We conclude by extending our results to tackle Nash equilibrium problems with even more general strategy profiles, where the nonnegativity constraints on realization plans are replaced with more general convex sets.
\end{abstract}

\keywords{sequential game; imcomplete information; perfect recall; sequence-form; Nash equilibrium; primal-dual algoithm; convex-optimization}

@@

\subsection{Some notation}
Let $m$ and $n$ be positive integers.
$\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ } z \geq 0\}$ is the nonnegative $n$-dimensional \textit{orthant}.  The components of a vector $z \in \mathbb{R}^n$ will be denoted $z_0$, $z_1$, ..., $z_{n-1}$ (indexing begins from $0$, and not $1$). The notation ``$z \ge 0$'' means that all the components of $z$ are nonnegative. For a vector $z = (z_0, z_1, ..., z_{n-1}) \in \mathbb{R}^n$, %% $\|z\|$ denotes the $2$-\textit{norm} of $z$ defined by $\|z\| := \sqrt{z^Tz} = \sqrt{\sum_{j=0}^{n - 1}{z_j^2}}$.
$(z)_+:=\text{max}(0, z) \in \mathbb{R}^{n}_+$ is the point-wise maximum of $z$ with $0$. For example, $((-2, \pi))_+ = (max(-2, 0), max(\pi, 0)) = (0, \pi)$.  The operator $(.)_+$ is the well-known (multi-dimensional) \textit{ramp} function.

$\mathbb{R}^{m \times n}$ denotes the space of all $m$-by-$n$ real matrices, and $0_{m,n}$ denotes the $m$-by-$n$ matrix of zeros. Given a matrix $A \in \mathbb{R}^{m \times n}$, its entries will be donoted $A_{i,j}$ so that
\[
A :=\left[
\begin{array}{cccc}
A_{0,0} & A_{0,1} & \cdots & A_{0,n-1} \\
A_{1,0} & A_{1,1} & \cdots & A_{1,n-1} \\
\vdots & \vdots & \ddots & \vdots\\
A_{m-1,0} & A_{m-1,1} & \cdots & A_{m-1,n-1}
\end{array}\right]
\]
The \textit{spectral norm} of a matrix $A$, denoted $\|A\|$, is defined to be the largest \textit{singular value} of $A$, i.e the largest \textit{eigen-value} of $A^TA$ (or equivalently, of $AA^T$).

\subsection{The problem}
The sequence-form for two-person zero-sum games was introduced in \cite{koller1992complexity}, and the theory was further developed in \cite{von1996efficient, vonequilibrium}, where it was established that for sequential two-person zero-sum games with imcomplete information and perfect recall, there exist sparse matrices $A \in \mathbb{R}^{n_1 \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are all linear in the size of the game tree (number of states in the game) and such that Nash equilibria correspond to pairs $(x, y)$ of \textit{realization plans} which solve the primal LCP (Linear Constrained Programming) problem
\begin{equation}
  \begin{aligned}
    & \underset{y,p}{\text{minimize}}
    & & p_0 \\
    & \text{subject to}
    &&-Ay + E_1^Tp \geq 0,\\
    &&&\hspace{4em}E_2y = e_2,\\
    &&&\hspace{5em} y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

and the dual LCP problem
\begin{equation}
  \begin{aligned}
    & \underset{x,q}{\text{maximize}}
    && -q_0 \\
    & \text{subject to}
    && -A^Tx - E_2^Tq \leq 0,\\
    &&&\hspace{4.5em} E_1x = e_1,\\
    &&&\hspace{5.5em} x \ge 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$ and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose entries are $-1$, $0$ or $1$, and each $e_k$ is a vector of the form $[1, 0, ..., 0]$. In this so-called \textit{sequence-form} representation, the strategy profile of player $k$ is the polyhedron
\begin{equation}
  Q_k := \{z \in \mathbb{R}^n_+ |\text{ }E_kz = e_k\}
\label{eq:polyhedron}
\end{equation}
Note that $p^Te_1 = p_0$ and $q^Te_2 = q_0$ by definition of $e_1$ and $e_2$. At a feasible point $(y, p, x, q)$, the \textit{duality gap} $\mathcal{G}(y, p, x, q)$ of these primal-dual pair of LCP problems is given by\footnote{The inequality being due to \textit{weak duality}.}
\begin{equation}
  0 \le \mathcal{G}(y, p, x, q) := p_0 - (-q_0) = p_0 + q_0
  \label{eq:dgap}
\end{equation}
Define the following closed convex sets
\begin{equation}
  \begin{aligned}
  \Sigma_1 &:= \{(y, p) \in \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}\text{ }|\text{ } -Ay + E_1^Tp \ge 0, E_2y = e_2, y \ge 0\}\\
  \Sigma_2 &:= \{(x, q) \in \mathbb{R}^{n_1} \times \mathbb{R}^{l_2}\text{ }|\text{ } -A^Tx - E_2^Tq \le 0, E_1x = e_1, x \ge 0\}
  \end{aligned}
  \label{eq:feasibility}
\end{equation}
It was shown (see Theorem 3.14 of \cite{vonequilibrium}) that a pair $(x, y)$ is a solution to the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} if (and only if) there exist vectors $p$ and $q$ such that $(y, p) \in \Sigma_1$ and $(x, q) \in \Sigma_2$. Thus the problem of finding a Nash equilibrium is equivalent to the problem of finding a point in the cartesian product $\Sigma_1 \times \Sigma_2$. Moreover, at equilibria, \textit{strong duality} holds and the value of the game equals $p_0 = -q_0$, i.e the duality gap $\mathcal{G}(y, p, x, q)$ defined in \eqref{eq:dgap} vanishes at equilibria.

@@

\begin{remark}  
It should be noted that for the the class of games considered here (sequential games with imcomplete information), the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger than what state-of-the-art LCP solvers can handle. See \cite{hoda2010smoothing}.
\end{remark}

Solving the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} exactly is impossible in practice \textbf{Cite Koller!!!}, and such a precision doesn't have any fundamental advantage. Instead, it is customary compute so-called Nash $\epsilon$-equlibria.

@@

\subsection{Outline of the article}
The next section features concepts from convex-analysis which will be used without further explanation. In section \ref{sec:related_work}, we give a brief overview of existing methods for computing Nash $\epsilon$-equilibria. In section \ref{sec:algo} we derive the proposed algorithm for producing a Nash $\epsilon$-equilibrium and prove its $\mathcal{O}(1/\epsilon)$ convergence rate. Section \ref{sec:results} presents practical results obtained both on simulated matrix games and the Kuhn 3-card Poker \cite{kuhn}.

\section{Useful convex-analysis}
\label{sec:notation}
We will need the following notations and concepts from convex-analysis in the sequel.
Given a \textit{convex subset} $C$ of $\mathbb{R}^n$, $i_C$ denotes its \textit{indicator function} defined by
\begin{equation}
  i_C(x) = \begin{cases}
    0, &\mbox{if } x \in C\\
    +\infty, &\mbox{otherwise}.
    \end{cases}
  \end{equation}
Note that $i_{C \cap D} = i_C + i_D$. The \textit{euclidean projection} onto $C$, denoted $\Pi_C$, is the function
$\Pi: \mathbb{R}^n \mapsto C$ which maps a point $x \in \mathbb{R}^n$ to the (necessarily unique) point $\Pi_C(x)$ of $C$ which is closest to $x$. Precisely,
\begin{equation}
  \Pi_C(x) := \underset{c \in C}{argimin}\text{ }\|c - x\|^2
\end{equation}
As an example, the euclidean projections onto the nonnegative orthant and the closed unit-disk $\bar{\Delta}$
% := \{x|\|x\| \le 1\}$
are $\Pi_{\mathbb{R}^n_+}(x) \equiv (x)_+$ and $\Pi_{\bar{\Delta}}(z) \equiv \dfrac{z}{max(\|z\|, 1)}$ respectively.

Let $f : \mathbb{R}^n \rightarrow (-\infty, +\infty]$ be a \textit{proper} (meaning that it is not identically $\infty$) convex function.
%% \textit{proper convex lower semi-continous function}
%% (\textit{p.c.l.s.c} for short).
The \textit{convex conjugate} of $f$ is the function $f^*: \mathbb{R}^n \rightarrow (-\infty, +\infty]$ defined by
\begin{equation}
  x \mapsto f^*(x) := \underset{z \in \mathbb{R}^n}{\text{max}}\text{ }z^Tx - f(z)
\end{equation}
For example, $i_{\mathbb{R}^{n}_+}^*(x) \equiv i_{\mathbb{R}^{n}_+}(-x)$; $(a^Tx)^* \equiv i_{\{0\}}(x - a)$; and $(\frac{1}{2}\|x\|^2)^* \equiv \frac{1}{2}\|x\|^2$. One easily verifies that $f^{**} = f$, if $f$ is \textit{lower semi-continuous}. The \textit{domain} of $f$ is defined by
\begin{equation}
dom f := \{x \in \mathbb{R}^n | f(x) < \infty\}
\end{equation}
If $dom f = \mathbb{R}^{n}$, then we say $f$ has \textit{full domain}.
Note that if $x \mapsto \dfrac{f(x)}{\|x\|}$ is unbounded, then $f^*$ has full domain.

Let $f$ be as above. Given $\tau > 0$, the \textit{proximal operator} of $f$, at rank
$\tau$, denoted $\text{prox}_{\tau f}$, is the function which maps a point $x \in \mathbb{R}^n$ to the (necessarily
unique) solution of the problem
\begin{equation}
  \underset{z \in \mathbb{R}^n}{argmin}\text{ }\frac{1}{2}\|z - x\|^2 + \tau f(z)
\end{equation}
It is easy to see that if $f$ is the indicator function of a convex set $C$, then $\text{prox}_{\tau f} = \Pi_C, \forall \tau > 0$. In this sense, proximal operators can be seen
as a generalization of euclidean projections.

The passionate reader should refer to \cite{rockafellar1997convex, combettes2011proximal} for a more elaborate exposition of convex-analytical concepts and their use in optimization.

@@

In \cite{hoda2010smoothing}, a nested iterative procedure using the Excessive Gap Technique (EGT) \cite{nesterov2005excessive} was used to to solve the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} in the following equivalent saddle-point form

@@

The primal-dual algorithm first developed in \cite{chambolle2010} was proposed \cite{chambolle2014ergodic} in to solve matrix games on simplexes. It should be stressed that such matrix games are considerably simpler than the games considered here. Indeed, the authors in \cite{chambolle2014ergodic} use the fact that computing the euclidean projection of a point unto a simplex can be done in linear time as in \cite{duchi2008efficient}. In contrast, no such efficient algorithm is known nor is likely to exist for the polyhedra $Q_k$ defined in \eqref{eq:polyhedron}, the strategy profiles for players in the games considered here. It should however be noted that such this projection can still be done iteratively using, for example, the algorithm in proposition 4.2 of \cite{combettes2010dualization}, for example. Unfortunately, the authors didn't provide a convergence rate for their algorithm, and as with any nested iterative scheme, one would have to solve this sub-problem with finer and finer precision.

Sampling techniques like the CFR (CounterFactual Regret minimization) and its many variants \cite{MartinZinkevichNIPS2007, lanctot2009monte, Bowling09012015} have also become state-of-the-art.

\section{Our main contribution}
\label{sec:algo}
We now present our main contribution, namely an $\mathcal{O}(1/\epsilon)$ primal-dual algorithm for computing Nash $\epsilon$-equilibria for sequential two-person zero-sum games with imcomplete information and perfect recall.
%% The algorithm is obtained from the general primal-dual Algorithm 39 proposed in \cite{chambolle2010}.

\subsection{Reformulation of the problem}
It should be noted that the difficulty of applying the general primal-dual Algorithm 39 proposed in \cite{chambolle2010} directly to problem \eqref{eq:primal_pb} lies in the difficulty of effectively computing the euclidean projections $\Pi_{Q_k}$ (the proximal operators). Theorem \ref{thm:pd} provides a trick to effectively evade this difficulty, by re-writting the problem into an equivalent saddle-point convex-concave problem for which the associated proximal operators are simple.

\begin{theorem}
  The Nash equilibrium LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} can be re-written in the saddle-point convex-concove form

@@

  where the $(n_1 + l_2)$-by-$(n_2 + l_1)$ matrix $K$ is defined by
  \begin{equation}
    K :=
    \left[
      \begin{array}{c|c}
        A & -E_1^T \\ \hline
        E_2 & 0_{l_2, l_1}
      \end{array}
      \right]
\end{equation}

@@

\subsection{The proposed algorithm and its $\mathcal{O}(1/\epsilon)$ convergence}
According to Theorem \ref{thm:pd}, solving the Nash equilibrium LCP problems \eqref{eq:primal_pb}, \eqref{eq:dual_pb} is equivalent to solving the saddle-point problem \eqref{eq:unconstrained_pb}.
In order to apply Algorithm 39 of \cite{chambolle2010}, we need formulae for the proximal operators of
$G_1$ and $G_2$. Now, either of $G_1$ and $G_2$ is a separable sum of the indicator function of a convex set, whose prox is simply the euclidean projection onto the set,  and a linear transformation $z \mapsto a^Tz$ whose prox is simply $z \mapsto z - \tau a$. Thus,  one computes their respective proximal operators $\text{prox}_{\tau G_1} : \mathbb{R}^{n_2} \times \mathbb{R}^{l_1} \rightarrow \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}$, $\text{prox}_{\sigma G_2}: \mathbb{R}^{n_1} \times \mathbb{R}^{l_2} \rightarrow \mathbb{R}^{n_1} \times \mathbb{R}^{l_2}$ as
  \begin{equation}
    \text{prox}_{\tau G_1}(y, p) = ((y)_+, p - \tau e_1), \hspace{1em} \text{prox}_{\sigma G_2}(x, q) = ((x)_+, q - \sigma e_2)
  \end{equation}

Finally, applying Algorithm 39 of \cite{chambolle2010} to problem \eqref{eq:unconstrained_pb} we obtain Algorithm \ref{Tab:algo_simplified}. Its $\mathcal{O}(1/\epsilon)$ convergence rate in terms of the duality gap is immediate. Indeed,
%% \subsection{Special case: $C_k = \mathbb{R}^{n_k}_+$} As discussed in subsection \ref{subsec:example_games},
%% the Nash equilibrium problem for two-person zero-sum simultaneous games and two-person zero-sum sequential games with imcomplete
%% information and perfect recall admits the formulation \eqref{eq:primal_pb}, with $C_k = \mathbb{R}^{n_k}_+$ (coding for nonnegativity constraints).
%% In such situations, $\Pi_{C_k}(z) \equiv (z)_+$ as already mentioned in \ref{sec:notation}, and Algorithm \ref{Tab:algo} reduces to the simpler Algorithm \ref{Tab:algo_simplified}.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e}%,times}
%\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}

%\usepackage[section]{placeins}
\usepackage{subfigure}

\def\B#1{\bm{#1}}
%\def\B#1{\mathbf{#1}}
\def\trans{\mathsf{T}}

%\renewcommand{\labelitemi}{--}

\newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}

\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}{Remark}
\newtheorem{fact}{Fact}

\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\prox}{Prox}

@@

@@

\begin{abstract}
We present a simple primal-dual algorithm
%with low costs per iteration,
for computing Nash-equilibria in two-person zero-sum sequential games
with imcomplete information and perfect recall (like Texas Hold'em
Poker). This primal-dual algorithm derives from the general
primal-dual algorithm developed
% by Chambolle and Pock
in \cite{chambolle2010,chambolle2014ergodic}. It should be noted that
the latter primal-dual algorithm is not directly applicable to compute
Nash equilibria for the aforementioned games considered here since the
strategy profiles of players (polyhedra) are so complex that it is not
possible
% (for example, simplexes for simultaneous games, and intersections of
% hyperplanes and nonnegative orthants, for sequential games)
 to effectively compute % (i.e using non-iterative procedures, etc.)
euclidean
% (metric)%
 projections (the proximal operators) thereupon.
The core of our technique is to rewrite the equilibrium problem as an
equivalent saddle-point problem for which the proximal operators can
be computed analytically.
 %% dualize the linear constraints that form the player's strategy profile, thus obtainining an equivalent saddle-point problem which is amenable to the primal-dual scheme.
Our algorithm only performs basic iterations
(i.e iterations involving no
% only involving basic operations like matvec, clipping, etc., and no
calls to external first-order oracles, etc.).
Our algorithm is applicable to a broad class of two-person zero-sum
games including simultaneous games and sequential games with
imcomplete information and perfect recall. The applicability to the
latter kind of games is thanks to the sequence-form representation von
Stengel's \cite{von1996efficient}
%% Unlike the EGT (Excessive Gap Technique) recently used in
%% \cite{hoda2010smoothing, gilpinfirst}, our algorithm has a cheap
%% and constant cost per iteration.
As an application, we apply our algorithm to solve Kuhn 3-card Poker.
%% We conclude by extending our results to tackle Nash equilibrium problems with even more general strategy profiles, where the nonnegativity constraints on realization plans are replaced with more general convex sets.
\end{abstract}

\textbf{keywords}: sequential game; imcomplete information; perfect
recall; sequence-form; Nash equilibrium; primal-dual algoithm;
split-feasibility problem; convex-optimization

@@

\subsection{Some notation and terminology}
Let $m$ and $n$ be positive integers.
$\mathbb{R}^{n}_+ := \{z \in \mathbb{R}^{n}\text{ }|\text{ } z \geq
0\}$ is the nonnegative $n$-dimensional \textit{orthant}.  The
components of a vector $z \in \mathbb{R}^n$ will be denoted $z_0$,
$z_1$, ..., $z_{n-1}$ (indexing begins from $0$, and not $1$). The
notation ``$z \ge 0$'' means that all the components of $z$ are
nonnegative. For a vector $z = (z_0, z_1, ..., z_{n-1}) \in
\mathbb{R}^n$, %% $\|z\|$ denotes the $2$-\textit{norm} of $z$ defined
%% by $\|z\| := \sqrt{z^Tz} = \sqrt{\sum_{j=0}^{n - 1}{z_j^2}}$.
$(z)_+:=\text{max}(0, z) \in \mathbb{R}^{n}_+$ is the point-wise
maximum of $z$ with $0$.
%% For example, $((-2, \pi))_+ = (max(-2, 0),
%% max(\pi, 0)) = (0, \pi)$.
The operator $(.)_+$ is the well-known
(multi-dimensional) \textit{ramp} function.
$\mathbb{R}^{m \times n}$ denotes the space of all $m$-by-$n$ real
matrices, $I_m$ denotes the $m$-by-$m$ identity matrix, and $0_{m,n}$
denotes the $m$-by-$n$ matrix of zeros.

Given a matrix $A \in \mathbb{R}^{m \times n}$, its \textit{spectral
  norm}, denoted $\|A\|$, is
 defined to be the largest \textit{singular value} of $A$, i.e the
 largest \textit{eigen-value} of $A^TA$ (or equivalently, of $AA^T$).

Given a subset $C$ of $\mathbb{R}^n$,
$i_C$ denotes its \textit{indicator function} defined by $i_C(x) = 0$
if $ x \in C$; $i_c(x) = \infty$ otherwise.

Let $f : \mathbb{R}^n \rightarrow (-\infty, +\infty]$ be a
  \textit{proper} (meaning that it is not identically $\infty$) convex
  function.
%% \textit{proper convex lower semi-continous function}
%% (\textit{p.c.l.s.c} for short).
The \textit{convex conjugate} of $f$ is the function $f^*:
\mathbb{R}^n \rightarrow (-\infty, +\infty]$ defined by $f^*(x) \equiv
  \underset{z \in \mathbb{R}^n}{\text{max}}\text{}z^Tx - f(z)$.
Given $\tau > 0$, the \textit{proximal operator}
of $f$ at rank $\tau$, denoted $\prox_{\tau f}$, is the
function which maps a point $x \in \mathbb{R}^n$ to the (necessarily
unique) solution of the problem $\underset{z \in
  \mathbb{R}^n}{argmin}\text{ }\frac{1}{2}\|z - x\|^2 + \tau f(z)$.

The interested reader should refer to
\cite{rockafellar1997convex,combettes2011proximal} for a more
elaborate exposition of convex-analytical concepts and their use in
optimization.

\subsection{Statement of the problem}
The sequence-form for two-person zero-sum games was introduced in
\cite{koller1992complexity}, and the theory was further developed in
\cite{von1996efficient,vonequilibrium}, where it was established that
for sequential two-person zero-sum games with imcomplete information
and perfect recall, there exist sparse matrices $A \in \mathbb{R}^{n_1
  \times n_2}$, $E_1 \in \mathbb{R}^{l_1 \times n_1}$, $E_2 \in
\mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1},
e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are
all linear in the size of the game tree (number of states in the game)
and such that Nash equilibria correspond to pairs $(x, y)$ of
\textit{realization plans} which solve the primal LCP
\begin{equation}
  \begin{aligned}
     \text{minimize }p_0, \hspace{.5em}\text{ subject to }(y,p) \in
     \mathbb{R}^{n_2} \times \mathbb{R}^{l_1}, -Ay +
    E_1^Tp \geq 0, E_2y = e_2, y \ge 0.
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

and the dual LCP
\begin{equation}
  \begin{aligned}
    \text{maximize}-q_0,\hspace{.5em}\text{subject to }(y,q) \in
    \mathbb{R}^{n_1} \times \mathbb{R}^{l_2}, -A^Tx - E_2^Tq \leq 0,
    E_1x = e_1, x \ge 0.
  \end{aligned}
  \label{eq:dual_pb}
\end{equation}
The vectors $p = (p_0, p_1, ..., p_{l_2 - 1}) \in \mathbb{R}^{l_2}$
and $q = (q_0, q_1, ..., q_{l_1 - 1}) \in \mathbb{R}^{l_1}$ are dual
variables. 
$A$ is the \textit{payoff matrix} and each $E_k$ is a matrix whose
entries are $-1$, $0$ or $1$, and each $e_k$ is a vector of the form
$[1, 0, ..., 0]$. In this so-called \textit{sequence-form}
representation, the strategy profile of player $k$ is the polyhedron
\begin{equation}
  Q_k := \{z \in \mathbb{R}^n_+ |\text{ }E_kz = e_k\}
\label{eq:polyhedron}
\end{equation}
Note that $p^Te_1 = p_0$ and $q^Te_2 = q_0$ by definition of $e_1$ and
$e_2$. At a feasible point $(y, p, x, q)$, the \textit{duality gap}
$\mathcal{G}(y, p, x, q)$ of these primal-dual pair of LCPs is given
by\footnote{The inequality being due to \textit{weak duality}.}
\begin{equation}
  0 \le \mathcal{G}(y, p, x, q) := p_0 - (-q_0) = p_0 + q_0
  \label{eq:dgap}
\end{equation}

%\subsection{Characterization of equilibria}
It was shown (see Theorem 3.14 of \cite{vonequilibrium}) that a pair
$(x, y) \in Q_1 \times Q_2$ of realization plans is a solution to the
LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} (i.e is a Nash
equilibrium for the game)  if and only if there exist vectors $p$ and
$q$ such that
\begin{equation}
\hspace{.25em} -Ay + E_1^Tp \ge 0, \hspace{.5em}A^Tx + E_2^Tq \ge
0, \hspace{.25em} x^T(-Ay + E_1^Tp) = 0, \hspace{.25em} y^T(A^Tx +
E_2^Tq) = 0.
\label{eq:feasibility}
\end{equation}

Moreover, at equilibria, \textit{strong duality} holds and the value of the game equals $p_0 = -q_0$, i.e the duality gap $\mathcal{G}(y, p, x, q)$ defined in \eqref{eq:dgap} vanishes at equilibria.

@@

\begin{remark}  
It should be noted that for the the class of games considered here (sequential games with imcomplete information), the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} are exceedingly larger than what state-of-the-art LCP solvers can handle. See \cite{hoda2010smoothing}.
\end{remark}

Solving the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} exactly is impossible in practice \textbf{Cite Koller!!!}, and such a precision doesn't have any fundamental advantage. Instead, it is customary compute so-called Nash $\epsilon$-equlibria.

@@

\subsection{Outline of the article}
We develop a $\mathcal{O}(1/\epsilon)$ primal-dual algorithm for computing Nash
$\epsilon$-equilibria for sequential two-person zero-sum games with
imcomplete information and perfect recall.

@@

In \cite{hoda2010smoothing}, a nested iterative procedure using the Excessive Gap Technique (EGT) \cite{nesterov2005excessive} was used to to solve the LCPs \eqref{eq:primal_pb} and \eqref{eq:dual_pb} in the following equivalent saddle-point form

@@

The primal-dual algorithm first developed in \cite{chambolle2010} was
proposed \cite{chambolle2014ergodic} in to solve matrix games on
simplexes. It should be stressed that such matrix games are
considerably simpler than the games considered here. Indeed, the
authors in \cite{chambolle2014ergodic} use the fact that computing the
euclidean projection of a point unto a simplex can be done in linear
time as in \cite{duchi2008efficient}. In contrast, no such efficient
algorithm is known nor is likely to exist for the polyhedra $Q_k$
defined in \eqref{eq:polyhedron}, the strategy profiles for players in
the games considered here. It should however be noted that such this
projection can still be done iteratively using, for example, the
algorithm in proposition 4.2 of \cite{combettes2010dualization} or any
CQ algorithm (since this projection sub-problem, is, in fact a
\textit{split-feasibility problem}), for example. Unfortunately, as
with any nested iterative scheme, one would have to solve this
sub-problem with finer and finer precision.

Sampling techniques like the CFR (CounterFactual Regret minimization)
and its many variants
\cite{MartinZinkevichNIPS2007,lanctot2009monte,Bowling09012015} have
also become state-of-the-art.

\section{Our contributions}
\label{sec:algo}
\subsection{Primal-dual algorithm}
It should be noted that the difficulty of applying the general
primal-dual Algorithm 39 proposed in \cite{chambolle2010} directly to
problem \eqref{eq:primal_pb} lies in the difficulty of effectively
computing the euclidean projections $\proj_{Q_k}$ (which are the
proximal operators here). Theorem \ref{thm:pd} provides a trick to
effectively evade this difficulty, by re-writting the problem into an
equivalent saddle-point convex-concave problem for which the
associated proximal operators are simple.

\begin{theorem}
  The Nash equilibrium LCPs \eqref{eq:primal_pb} and
  \eqref{eq:dual_pb} are equivalent to the following convex-concave
  saddle-point problems

@@

  where the $(n_1 + l_2)$-by-$(n_2 + l_1)$ matrix $K$ is defined by
  \begin{equation}
    K :=
    \left[
      \begin{array}{cc}
        A & -E_1^T \\
        E_2 & 0_{l_2, l_1}
      \end{array}
      \right]
\end{equation}

@@

% \subsection{The proposed algorithm and its $\mathcal{O}(1/\epsilon)$
% convergence}

\paragraph{In summary.}
According to Theorem \ref{thm:pd}, solving the Nash equilibrium LCPs
\eqref{eq:primal_pb}, \eqref{eq:dual_pb} is equivalent to solving the
saddle-point problem \eqref{eq:unconstrained_pb}.
In order to apply Algorithm 39 of \cite{chambolle2010}, we need
formulae for the proximal operators of
$G_1$ and $G_2$. Now, either of $G_1$ and $G_2$ is a separable sum of
the indicator function of a convex set, whose prox is simply the
euclidean projection onto the set,  and a linear transformation $z
\mapsto a^Tz$ whose prox is simply $z \mapsto z - \tau a$. Thus,  one
computes their respective proximal operators $\prox_{\tau G_1} :
\mathbb{R}^{n_2} \times \mathbb{R}^{l_1} \rightarrow \mathbb{R}^{n_2}
\times \mathbb{R}^{l_1}$, $\prox_{\sigma G_2}: \mathbb{R}^{n_1} \times
\mathbb{R}^{l_2} \rightarrow \mathbb{R}^{n_1} \times \mathbb{R}^{l_2}$
as

  \begin{equation}
    \prox_{\tau G_1}(y, p) = ((y)_+, p - \tau e_1), \hspace{1em}
    \prox_{\sigma G_2}(x, q) = ((x)_+, q - \sigma e_2)
  \end{equation}

Finally, applying Algorithm 39 of \cite{chambolle2010} to problem \eqref{eq:unconstrained_pb} we obtain Algorithm \ref{Tab:algo_simplified}. Its $\mathcal{O}(1/\epsilon)$ convergence rate in terms of the duality gap is immediate. Indeed,
%% \subsection{Special case: $C_k = \mathbb{R}^{n_k}_+$} As discussed in subsection \ref{subsec:example_games},
%% the Nash equilibrium problem for two-person zero-sum simultaneous games and two-person zero-sum sequential games with imcomplete
%% information and perfect recall admits the formulation \eqref{eq:primal_pb}, with $C_k = \mathbb{R}^{n_k}_+$ (coding for nonnegativity constraints).
%% In such situations, $\proj_{C_k}(z) \equiv (z)_+$ as already mentioned in \ref{sec:notation}, and Algorithm \ref{Tab:algo} reduces to the simpler Algorithm \ref{Tab:algo_simplified}.
